---
title: "Introduction to R"
author: "Nathaniel Tomasetti"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

# What is R?

--
R is a programming language

-It works with **objects** and **functions**.

-It can be made to do almost any mathematical operation to data. 

-It easily allows people to create their own functions and share them with the world.

-Its basic functions are similar to what you can do in Excel.

-Its advanced functions make it both extremely powerful and widely used.

-It is free and available on any operating system.

---

class:middle, inverse, center
# Objects in R

---

Objects allow things to be created and stored

--

They can be variables
```{r}
x = 5
x + 1
```
--

They can be vectors, or matrices
```{r}
y = c(1, 4, 2, 6, 5, 0)
z = matrix(y, nrow = 2, ncol = 3)
y
z
```
--

An object can even by an entire dataset!

---

Datasets are usually too big to look at all at once. But we can look at parts of it easily

```{r}
head(mtcars)
```

--

Or a summary, or the structure
```{r, eval=FALSE}
summary(mtcars)
str(mtcars)
```

---
class:middle, inverse, center
#Functions in R

---

Functions are a set of instructions to turn an input into an output

```{r}
myMean = function(x){
  N = length(x)
  mean = sum(x) / N
  return(mean)
}
x = c(4, 1, 6, -3, 7, 1)
myMean(x)
```

--

It's easy to give them multiple inputs

```{r}
addAndDouble = function(x, y){
  2 * (x + y)
}
addAndDouble(5, 2)
```

---

R comes with many built in functions
```{r}
mean(x)
```
--
```{r}
median(x)
```
--
```{r}
sd(x)
```
--
```{r}
var(x)
```
--
```{r}
cor(x, c(0, -5, 1, 3, 11, 2))
```

---

R can do all of the regression you would do in eViews, Excel, Stata, SAS etc.

```{r}
x = rnorm(1000)
y = 2 + 3 * x + rnorm(1000, sd = 0.1)
linReg = lm(y ~ x)
summary(linReg)
```

---

If R can't do something, someone has probably written a function to do it.

```{r, eval=FALSE}
ggplot() + geom_point(aes(x, y))
```

The function ggplot() is part of the ggplot2 package. Install it once per computer

```{r, eval=FALSE}
installed.packages("ggplot2")
```

Then use library(ggplot2) in your code and R can use ggplot2 functions until the next time you start R.

```{r, eval=FALSE}
library(ggplot2)
ggplot() + geom_point(aes(x, y))
```

---

class: middle, center
```{r, echo=FALSE}
library(ggplot2)
ggplot() + geom_point(aes(x, y))
```

---

If you're ever stuck on how to use a function, you can get help with ?

```{r, eval=FALSE}
?mean
?ggplot
```


-- 

There's also tonnes of online resources available,


- https://stackoverflow.com/ is a great resource for programming help

- https://cran.r-project.org/ has documentation on almost every function

- https://www.datacamp.com/ has nice interactive courses on learning R.

